---
title: "P8105_HW1"
author: "Haley Zylberberg"
output: github_document
---

#Problem 1: This problem focuses the use of inline R code, plotting, and the behavior of ggplot for variables of different types.

To start, download the penguins dataset in the console directly.

Next, call in moderndrive and tidyverse library.

```{r setup}
library(moderndive)
library(tidyverse)
```

Load early_january_weather dataset.
```{r dataset}
data("early_january_weather")
```

Examine the dataset using the help and View functions.

```{r dataset help and view, eval = FALSE}
help(early_january_weather)
View(early_january_weather)
```

The dataset 'Early January Weather' depicts the hourly meterological data for the airport `r unique(early_january_weather$origin)` for the month of January 2013. There are  `r nrow(early_january_weather)` rows and `r ncol(early_january_weather)` columns (i.e. variables) in the dataset. The variable names are `r names(early_january_weather)`. 

* The following are summary statistics:
  * The mean temperature in F is `r mean(early_january_weather$temp)` (SD `r sd(early_january_weather$temp)`), with minimum value `r min(early_january_weather$temp)` and maximum value `r max(early_january_weather$temp)`. 
  * The mean dewpoint in F is `r mean(early_january_weather$dewp)` ( SD `r sd(early_january_weather$dewp)`), with minimum value `r min(early_january_weather$dewp)` and maximum value `r max(early_january_weather$dewp)`. 
  * The mean humidity is `r mean(early_january_weather$humid)` (SD `r sd(early_january_weather$humid)`), with minimum value `r min(early_january_weather$humid)` and maximum value `r max(early_january_weather$humid)`. 
  * The mean wind speed in mph is `r mean(early_january_weather$wind_speed)` (SD `r sd(early_january_weather$wind_speed)`), with minimum value `r min(early_january_weather$wind_speed)` and maximum value `r max(early_january_weather$wind_speed)`. 
  * The mean precipitation in inches is `r mean(early_january_weather$precip)` (SD `r sd(early_january_weather$precip)`), with minimum value `r min(early_january_weather$precip)` and maximum value `r max(early_january_weather$precip)`. 
  * The mean visibility in miles is `r mean(early_january_weather$visib)` (SD `r sd(early_january_weather$visib)`), with minimum value `r min(early_january_weather$visib)` and maximum value `r max(early_january_weather$visib)`. 
  * Statistics could not be obtained for the variables wind gust, wind direction, and pressure due to missing variables (`r sum(is.na(early_january_weather$wind_gust))` missing values for wind gust, `r sum(is.na(early_january_weather$wind_dir))` missing values for wind direction, and `r sum(is.na(early_january_weather$pressure))` missing values for pressure). 
  
Next make a scatterplot of temp (y) vs time (hour) adding in color points with the humid variable. 

```{r yx_scatter}
ggplot(early_january_weather, aes(x = hour, y = temp, 
color = humid)) + geom_point()
```

