---
title: "P8105_HW1"
author: "Haley Zylberberg, hmz2105"
output: github_document
---

# Problem 1: This problem focuses the use of inline R code, plotting, and the behavior of ggplot for variables of different types.

To start, download the penguins dataset in the console directly.

Next, call in moderndrive and tidyverse library.

```{r setup}
library(moderndive)
library(tidyverse)
```

Load early_january_weather dataset.
```{r dataset}
data("early_january_weather")
```

Examine the dataset using the help and View functions.

```{r dataset help and view, eval = FALSE}
help(early_january_weather)
View(early_january_weather)
```

The dataset 'Early January Weather' depicts the hourly meterological data for the airport `r unique(early_january_weather$origin)` for the month of January 2013. There are  `r nrow(early_january_weather)` rows and `r ncol(early_january_weather)` columns (i.e. variables) in the dataset. The variable names are `r names(early_january_weather)`. 

* The following are summary statistics:
  * The mean temperature in F is `r mean(early_january_weather$temp)` (SD `r sd(early_january_weather$temp)`), with minimum value `r min(early_january_weather$temp)` and maximum value `r max(early_january_weather$temp)`. 
  * The mean dewpoint in F is `r mean(early_january_weather$dewp)` ( SD `r sd(early_january_weather$dewp)`), with minimum value `r min(early_january_weather$dewp)` and maximum value `r max(early_january_weather$dewp)`. 
  * The mean humidity is `r mean(early_january_weather$humid)` (SD `r sd(early_january_weather$humid)`), with minimum value `r min(early_january_weather$humid)` and maximum value `r max(early_january_weather$humid)`. 
  * The mean wind speed in mph is `r mean(early_january_weather$wind_speed)` (SD `r sd(early_january_weather$wind_speed)`), with minimum value `r min(early_january_weather$wind_speed)` and maximum value `r max(early_january_weather$wind_speed)`. 
  * The mean precipitation in inches is `r mean(early_january_weather$precip)` (SD `r sd(early_january_weather$precip)`), with minimum value `r min(early_january_weather$precip)` and maximum value `r max(early_january_weather$precip)`. 
  * The mean visibility in miles is `r mean(early_january_weather$visib)` (SD `r sd(early_january_weather$visib)`), with minimum value `r min(early_january_weather$visib)` and maximum value `r max(early_january_weather$visib)`. 
  * Statistics could not be obtained for the variables wind gust, wind direction, and pressure due to missing variables (`r sum(is.na(early_january_weather$wind_gust))` missing values for wind gust, `r sum(is.na(early_january_weather$wind_dir))` missing values for wind direction, and `r sum(is.na(early_january_weather$pressure))` missing values for pressure). 
  
Next make a scatterplot of temp (y) vs time (hour) adding in color points with the humid variable. 

```{r yx_scatter}
ggplot(early_january_weather, aes(x = hour, y = temp, 
color = humid)) + geom_point()
```

Next save plot.

```{r save}
ggsave("scatter_plot_HW1.pdf", height = 4, width = 6)
```

This scatterplot shows that

# Problem 2: This problem is intended to emphasize variable types and introduce coercion; some awareness of how R treats numeric, character, and factor variables is necessary for working with these data types in practice.

* Will create a data frame comprised of:
  * a random sample of size 10 from a standard Normal distribution
  * a logical vector indicating whether elements of the sample are greater than 0
  * a character vector of length 10
  * a factor vector of length 10, with 3 different factor “levels”

```{r HW1_df_create and print}

set.seed(123)

HW1_df = tibble(
  random_sample = rnorm(10),
  random_sample_pos = random_sample > 0,
  random_sample_char = c("A", "B", "C", "D", "E", "F", "G", "H", "I", "J"),
  random_sample_fac = factor(rep(c("Level1", "Level2", "Level3"), length.out = 10))
)

print(HW1_df)
```

Calculate mean of each variable.

```{r pull variables and calculate mean}
mean_random_sample = mean(pull(HW1_df, random_sample))
mean_random_sample_pos = mean(pull(HW1_df, random_sample_pos))
mean_random_sample_char = mean(pull(HW1_df, random_sample_char))
mean_random_sample_fac = mean(pull(HW1_df, random_sample_fac))
```

The mean of the variable random sample is `r mean(mean_random_sample)`. The mean of the logical vector is `r mean(mean_random_sample_pos)`. We cannot calculate the mean of a character or factor variable as they are not numeric variables. 

Convert logical, character, and factor vectors to numeric.

```{r}
HW1_df$random_sample_pos <- as.numeric(HW1_df$random_sample_pos)
HW1_df$random_sample_char <- as.numeric(HW1_df$random_sample_char)
HW1_df$random_sample_fac <- as.numeric(HW1_df$random_sample_fac)
```
